Selective electroporation of the viral envelope

### 2021-04-29 documents/pure_electroporation.tex
Unicode math notetaking?!

@@

While both mechanisms hinge on time constants, those discussed in the previous paper referred to the time for viscous loss of energy into the surrounding medium and the normal modes of the viral envelope; here the time constants relate to capacitive charging, and Smoluchowski processes. The relevant damage quantities were mechanical displacement and membrane stress, whereas here we consider transmembrane voltage and pore diameter.

At the risk of self-plagarism, we will restate some precedent

@@

It is, in general, something of a singular outlier in that it is one of very few biological mechanisms relevant at practical electric field intensities. Along with thermal treatment, it also performs an operation something like integration, distinct from absolute-value. It is perhaps this property which provides a small degree of tunability.

\footnote{It is amusing to draw parallels between electroporation and Bialke et al's reports on spacecraft reaction wheels. In both cases, destructive process are charging of the external shell (the spacecraft itself), a highly resistive layer (a layer of grease on the ball bearings), and a }

Varactor-like parametric, as the bilayer is squished by the potential;

@@

These regimes of intensity may possibly be accessible via ultrashort optical techniques. Similarly, Tsen \cite{Studies2014} report a non-electroporative mechanism, but these results have not been replicated at that intensity\cite{No2011}.

@@

Electroporation\cite{Electroporation1988}, now a common laboratory procedure, occurs when an enormous electric field causes ions in solution to diffuse across a cell membrane, capacitively charging it, provoking a change in lipid bilayer conformation\cite{Membrane2016}, reversibly rendering the membrane permeable. This is one among many other, more subtle mechanisms; the minutia of poration can become extraordinarily complex in certain circumstances and are beyond the author's current understanding\cite{Theoretical2007}. Hydrodynamic

@@

\subsection*{Basic scaling laws}

The most obvious difference is scale. Viruses are minute. All other properties of the, the classic Schwan equation implies that the steady-state DC transmembrane voltage is proportional to the size of the particle.

Madiyar's Vaccinia electroporation data aligns with. Assuming that the host porates at 0.05 MV/m, the 250 nm vaccinia might be expected to have a threshold of 4 MV/m.

However, this does not explain

@@

Some extremely simplified scaling laws for membrane charging effects due to short (sub Tc) pulses are given by Schoenbach \cite{Bioelectric2007}. This is based primarily on an equivalent circuit model, we largely neglect the electrodynamics that apply with short pulses and also the conductivity change due to pore formation. Electroporation is a highly nonlinear effect\cite{Letter1974}, and this .

@@

@@

want to more accurately represent the - most importantly, the core permittivity seems like it should alter the reflection significantly at high frequencies, but isn't usually treated

@@

Basic dry-weight measurements seem to suggest that influenza contains about 60\% water by weight \cite{lauffer1944biophysical}. Has diffusion coefficients and maintains a Donnan ion size equilibrium similar to that of lipid bilayers \cite{Effect2015b}.

@@

\subsection{Electrodynamic considerations}

@@

\subsection{Optimal pulse}

Differential

This bears considerable similarity to e.g. radar waveform design and compressed sensing, and also problems of coherent control in quantum systems. In many cases, the optimal control is equal to the transfer function

@@

"Underactuated";

@@

\subsection{direct method Pseudospectral solvers}

@@

First the objective function with a constant integral control power constraint was used, but this did not yield

\subsection{Optimal control solvers}

>>>>

\usepackage{physics}
\usepackage{bm}
\newcommand{\bstar}[1]{ {#1}^{\bm*} }

@@

While both mechanisms hinge on a set of time constants, those discussed in the previous paper referred to the time for viscous loss of energy into the surrounding medium and the normal modes of the viral envelope; here the time constants relate to capacitive charging, and Smoluchowski processes. Previously, the relevant quantities describing damage to the virion were mechanical displacement and membrane stress, whereas here we consider transmembrane voltage and pore diameter.

At the risk of self-plagarism, we will restate some precedent.

@@

It is, in general, something of a singular outlier in that it is one of very few biological mechanisms relevant at practical electric field intensities. Similar to thermal treatment, it also performs an operation something like integration with a threshold, a process distinct from normal absolute-value. It is perhaps this property which provides a small degree of tunability.

\footnote{It is amusing to draw parallels between electroporation and Bialke et al's reports on spacecraft reaction wheels. In both cases, destructive process involve differential charging of the external shell (the spacecraft itself) against an internal shell (the reaction wheels), producing a potential difference across a highly resistive layer (a layer of grease on the ball bearings)}

Varactor-like parametric properties as the bilayer is squished by the potential. hydrodynamic

The same Debye screening effects are also of some relevance.

@@

These regimes of intensity may possibly be accessible via ultrashort optical techniques. Similarly, Tsen \cite{Studies2014} report a non-electroporative mechanism, but these results have not been successfully independently replicated at that intensity\cite{No2011}.

@@

Electroporation\cite{Electroporation1988}, now a common laboratory procedure, occurs when an enormous electric field causes ions in solution to diffuse across a cell membrane, capacitively charging it, provoking a change in lipid bilayer conformation\cite{Membrane2016}, either reversibly rendering the membrane permeable or, if the number and size of pores is considerable, the membrane may be permanently destroyed. This is one among many other, more subtle mechanisms; the minutia of poration can become extraordinarily complex in certain circumstances and are beyond the author's current understanding\cite{Theoretical2007}. Hydrodynamic

@@

\section*{Basic scaling laws}

The most obvious difference is scale. Viruses are minute. All other properties of the system being equal, the standard Schwan equation for transmembrane voltage implies that the steady-state DC voltage is proportional to the size of the particle. In the case of a 100 nm virus

Madiyar's Vaccinia electroporation data aligns with. Assuming that the host porates at 0.05 MV/m, the 250 nm vaccinia might be expected to have a threshold of 4 MV/m. Mutatoyo's data also agree.

This does not agree with the two reports. It does not seem plausible that the

@@

Some extremely simplified scaling laws for membrane charging effects due to short (sub-first time constant) pulses are given by Schoenbach \cite{Bioelectric2007}. This is based primarily on an equivalent circuit model, we largely neglect the electrodynamics that apply with short pulses and also the conductivity change due to pore formation, modelled by [insert equation from "ADC" paper]. Electroporation is a highly nonlinear effect\cite{Letter1974}, and this .

@@

Regrettably, despite their diminutive radii, viruses and host cells may well have overlapping time constants due to the other dielectric properties. However, even in the case where both time constants are nearly equivalent, a marked discrepancy in the time evolution of short pulses can be seen.

[step response]

And so, despite having similar time constants, a short gaussian pulse [] long

@@

Equivalent circuit models want to more accurately represent the - most importantly, the core permittivity intuitively seems like it should alter the reflection significantly at high frequencies, but isn't usually treated

@@

Dry-weight measurements seem to suggest that influenza contains about 60\% water by weight \cite{lauffer1944biophysical}. Has diffusion coefficients and maintains a Donnan ion size equilibrium similar to that of lipid bilayers \cite{Effect2015b}.

@@

\subsection*{Electrodynamic considerations}

@@

\subsection*{Optimal pulse}

We must apologize once again to the mathematicians for sins of terminology.

This bears considerable similarity to e.g. radar waveform design and compressed sensing; problems of coherent control of quantum systems. In many cases, the optimal control is equal to the transfer function

@@

"Underactuated" robotics

@@

\subsection*{direct method Pseudospectral solvers}

@@

An objective function with a constant integral control power constraint, $t_{end} = \int_0^t_{end} u(t)^2 dt$ was used, but this did not appear to yield meaningful results.

\subsection*{Optimal control solvers}

\subsection*{Nondimensionalization}

The coefficients have a very small value; even time has a very small value. This was believed to account for at least some of the

$$\dv[2]{x_V}{t} = 
\frac{R a_{1V}}{b_{1V}} \dv[2]{u}{t} 
+ \frac{R a_{2V}}{b_{1V}}  \dv{u}{t} 
+ \frac{R a_{3V}}{b_{1V}}  u
- \frac{b_{2V}}{b_{1V}} \dv{x_V}{t}
- \frac{b_{3V}}{b_{1V}} x_V $$

$$\dv[2]{x_H}{t} = 
\frac{R a_{1H}}{b_{1H}} \dv[2]{u}{t} 
+ \frac{R a_{2H}}{b_{1H}}  \dv{u}{t} 
+ \frac{R a_{3H}}{b_{1H}}  u
- \frac{b_{2H}}{b_{1H}} \dv{x_H}{t}
- \frac{b_{3H}}{b_{1H}} x_H $$

%x2  = (R*a1*u2 + R*a2*u1 + R*a3*u0 - b2*x1 - b3*x0)/b1
% https://math.stackexchange.com/questions/3435173/how-to-nondimensionalize-a-second-order-differential-equation

let $x_V^{\bm*} = \frac{x_V}{x_{VK}}$ and $t^{\bm*} = \frac{t}{t_{K}}$ where $_K$ denotes a nondimensionalization constant ($t_0$ is switched to $t_K$ to avoid confusion with x0 used in simulation). (the forcing functional u is dependent on t, but this is implied).\\ \\

$$\frac{d^n}{dt^n} = \frac{1}{{t_K}^n}\frac{d^n}{d\bstar{t}^n}$$\\

$$\dv[2]{x_V}{t} = 
\frac{R a_{1V}}{b_{1V}} \dv[2]{u}{t} 
+ \frac{R a_{2V}}{b_{1V}}  \dv{u}{t} 
+ \frac{R a_{3V}}{b_{1V}}  u
- \frac{b_{2V}}{b_{1V}} \dv{x_V}{t}
- \frac{b_{3V}}{b_{1V}} x_V $$

GEKKO's .dt() operator allows u0 to be chosen as the control. 
Otherwise, the DAE form requires u2 to be used as the control.

denormalize u?
